Study 1 Analysis

### 2024-10-31 results.qmd
outlier checks and some model updates

s1_agg <- s1 |> 
  filter(appliance !="Total kWh") |> 
  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), 
            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif)) |> 
  mutate(matched_goal = (pct_change == pct_goal))

@@

mted1 <- as.data.frame(describe_posterior(s1_bb1, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mted1) <- c("Term", "Estimate"," 95% CrI Lower", " 95% CrI Upper", "pd")

mted1 |>
  mutate(across(c("Estimate", " 95% CrI Lower", " 95% CrI Upper"), 
                \(x) plogis(x))) |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |>
  tibble::remove_rownames() |>
  mutate(Term = stringr::str_remove(Term, "b_")) |>
  kable(booktabs = TRUE)

@@

@@

posterior_predict(s1_bb1,ndraws=200) |> array_branch(margin=1) |> 
  map_dfr(
    function(yrep_iter){
      s1_agg4 %>% ungroup() |>  mutate(pred=yrep_iter)
    },
    .id='iter'
  ) |> mutate(iter=as.numeric(iter)) |> 
  ggplot(aes(pred, group=iter)) +
  # Scale density to match the count scale
  geom_line(stat='density', aes(y=..count..), 
            alpha=0.7, color='blue') +
  geom_bar(data=s1_agg4, aes(mg, fill=refClass),
           inherit.aes=FALSE) +
  facet_wrap(~refClass)
```

@@

s2_bb2_r_g <- brm(
  mg | trials(2) ~ refClass*rounded*pct_goal,
  family = beta_binomial(),
  data = s2_agg2,
  cores=4,
  iter=5000,
  file=paste0(here::here("data/model_cache",'s2_bb2_r_g.rds'))
)

mted2 <- as.data.frame(describe_posterior(s2_bb2_r_g, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mted2) <- c("Term", "Estimate"," 95% CrI Lower", " 95% CrI Upper", "pd")

mted2 |>
  mutate(across(c("Estimate", " 95% CrI Lower", " 95% CrI Upper"), 
                \(x) plogis(x))) |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |>
  tibble::remove_rownames() |>
  mutate(Term = stringr::str_remove(Term, "b_")) |>
  kable(booktabs = TRUE)
```

@@

>>>>

s1_agg <- s1 |> 
  filter(appliance !="Total kWh") |> 
  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), 
            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
            n_change = sum(value!=family),
            state_p_dif=mean(state_p_dif),
            state_f_dif=mean(state_f_dif),
            n_less_avg = sum(less_avg)) |> 
  mutate(matched_goal = (pct_change == pct_goal), 
             error = pct_change - pct_goal,
            abs_error = abs(error))

@@

mted1 <- as.data.frame(describe_posterior(s1_bb1, centrality = "Median"))[, c(1,2,4,5,6)]
#colnames(mted1) <- c("Term", "Estimate"," 95% CrI Lower", " 95% CrI Upper", "pd")
colnames(mted1) <- c("Term", "Estimate", "95% CrI Lower", "95% CrI Upper", "pd")

intercept <- mted1$Estimate[mted1$Term == "b_Intercept"]

# mted1 |>
#   mutate(
#     Estimate_prob = plogis(Estimate + if_else(Term == "b_Intercept", 0, intercept)),
#     Lower_prob = plogis(`95% CrI Lower` + if_else(Term == "b_Intercept", 0, intercept)),
#     Upper_prob = plogis(`95% CrI Upper` + if_else(Term == "b_Intercept", 0, intercept))
#   )

t1 <- mted1 |>
  mutate(Term = stringr::str_remove(Term, "b_")) |>
  mutate(across(c("Estimate", "95% CrI Lower", "95% CrI Upper"), 
                \(x) if_else(Term == "Intercept", plogis(x), plogis(x + intercept)))) |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |>
  mutate(Term = if_else(Term == "Intercept", "Intercept (kWh)", Term)) |>
  tibble::remove_rownames() |>
  kable(booktabs = TRUE)

# plogis(-.58+ -.70) = .22

# parameters::parameters(s1_bb1, effect = "fixed")

@@

::: {#tbl-s1-bb}
|Term               | Estimate| 95% CrI Lower| 95% CrI Upper| pd|
|:------------------|--------:|-------------:|-------------:|--:|
|Intercept (kWh)    |     0.36|          0.35|          0.36|  1|
|refClassPercentage |     0.22|          0.21|          0.22|  1|
|refClassUSD        |     0.09|          0.09|          0.09|  1|

: **Experiment 1 main effect**. Beta binomial regression results for the main effect of Reference Class.

:::

@@

posterior_predict(s1_bb1,ndraws=200) |> array_branch(margin=1) |> 
  map_dfr(
    function(yrep_iter){
      s1_agg4 %>% ungroup() |>  mutate(pred=yrep_iter)
    },
    .id='iter'
  ) |> mutate(iter=as.numeric(iter)) |> 
  ggplot(aes(pred, group=iter)) +
  # Scale density to match the count scale
  geom_line(stat='density', aes(y=..count..), 
            alpha=0.7, color='blue') +
  geom_bar(data=s1_agg4, aes(mg, fill=refClass),
           inherit.aes=FALSE) +
  facet_wrap(~refClass)

s1_bb1 %>%
  gather_draws(b_Intercept, b_refClassPercentage, b_refClassUSD) %>%
   mutate(Term = stringr::str_remove(.variable, "b_")) |>
   mutate(prob=plogis(.value)) |>
  mutate(Term = if_else(Term == "Intercept", "kWh", Term)) |>
  ggplot(aes(x = prob, fill = Term)) +
  geom_density(alpha = 0.6) +
    stat_halfeye() +
  theme_minimal() +
  labs(title = "Posterior Distributions of Reference Class Estimates",
       x = "Coefficient Value",
       fill = "Reference Class")

@@

s2_bb2_r_g <- brm(
  mg | trials(2) ~ refClass*rounded*pct_goal,
  family = beta_binomial(),
  data = s2_agg2,
  cores=4,
  iter=5000,
  file=paste0(here::here("data/model_cache",'s2_bb2_r_g.rds'))
)
mted2 <- as.data.frame(describe_posterior(s2_bb2_r_g, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mted2) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

# mted2 |>
#   mutate(across(c("Estimate", " 95% CrI Lower", " 95% CrI Upper"), 
#                 \(x) plogis(x))) |>
#   mutate(across(where(is.numeric), \(x) round(x, 3))) |>
#   tibble::remove_rownames() |>
#   mutate(Term = stringr::str_remove(Term, "b_")) |>
#   kable(booktabs = TRUE)

 mted2 |>
  mutate(Term = stringr::str_remove(Term, "b_")) |>
  mutate(across(c("Estimate", "95% CrI Lower", "95% CrI Upper"), 
                \(x) if_else(Term == "Intercept", plogis(x), plogis(x + intercept)))) |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |>
  mutate(Term = if_else(Term == "Intercept", "Intercept (kWh)", Term)) |>
  tibble::remove_rownames() |>
  kable(booktabs = TRUE)

@@

s2_bb2_r_g |> emmeans( ~refClass, type="response") |>
  gather_emmeans_draws() |>
 condEffects(refClass) + labs(y="Absolute Deviation From Band", x="Band Type")

### 2024-10-31 study1.qmd
outlier checks and some model updates

  df_long <- df %>%
  select(-matches("01$|02$|03$|04$|05$|06$")) %>%
    pivot_longer(
      cols = all_of(action_plan_cols),
      names_to = c("modality", "state", "item_num"),
      names_pattern = "^([ADP])\\d+(CAL|MASS|COL|TEX)(\\d+)$",
      values_to = "value"
    ) |> select(id,state1,state2,refClass,calc,edu,modality,state,item_num,value,id2)

@@

s1_agg <- s1 |> 
  filter(appliance !="Total kWh") |> 
  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), 
            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
            n_change = sum(value!=family),
            state_p_dif=mean(state_p_dif),
            state_f_dif=mean(state_f_dif),
            n_less_avg = sum(less_avg)) |> 
  mutate(matched_goal = (pct_change == pct_goal))

k <- s1_agg |> group_by(id,pct_goal) |>  summarize(total_kWh=sum(total_kWh),orig_kWh=sum(orig_kWh),pct_change=round((orig_kWh-total_kWh)/orig_kWh,3)) |> 
  mutate(matched_goal = (pct_change == pct_goal))

@@

@@

# Step 2: Summarize the percentage of participants meeting the goal for each condition
summary_pct <- s1_agg |>
  group_by(refClass) |>
  summarise(
    matched_count = sum(matched_goal),
    total_count = n(),
    pct_matched = matched_count / total_count * 100,
    .groups = "drop"
  )

@@

>>>>

  df_long <- df %>%
  select(-matches("01$|02$|03$|04$|05$|06$")) %>%
    pivot_longer(
      cols = all_of(action_plan_cols),
      names_to = c("modality", "state", "item_num"),
      names_pattern = "^([ADP])\\d+(CAL|MASS|COL|TEX)(\\d+)$",
      values_to = "value"
    ) |> select(id,state1,state2,refClass,calc,edu,modality,state,item_num,value,id2,Duration__in_seconds_)

@@

s1_agg <- s1 |> 
  filter(appliance !="Total kWh") |> 
  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), 
            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
            n_change = sum(value!=family),
            state_p_dif=mean(state_p_dif),
            state_f_dif=mean(state_f_dif),
            n_less_avg = sum(less_avg),
            duration=first(Duration__in_seconds_)) |> 
  mutate(matched_goal = (pct_change == pct_goal), 
             error = pct_change - pct_goal,
            abs_error = abs(error),
            close_match = abs_error <= 0.02)

# s1_agg |> filter(matched_goal==FALSE, abs_error<.02) |> pull(abs_error)

@@

ggplot(s1_agg |> filter(abs_error<1), aes(x = refClass, y = abs_error, fill = refClass)) +
    geom_boxplot() +
    labs(
        title = "Absolute Error by Reference Class",
        x = "Reference Class",
        y = "Absolute Error"
    ) +
    theme_minimal()

ggplot(s1_agg |> filter(abs_error<1), aes(x = abs_error, fill = refClass)) +
    geom_histogram(binwidth = 0.05, alpha = 0.6, position = "identity") +
    labs(
        title = "Distribution of Absolute Error by Reference Class",
        x = "Absolute Error",
        y = "Frequency"
    ) +
    theme_minimal() + facet_wrap(~refClass)
```

@@

# Step 2: Summarize the percentage of participants meeting the goal for each condition
summary_pct <- s1_agg |>
  group_by(refClass) |>
  summarise(
    matched_count = sum(matched_goal),
    matched_count2 = sum(close_match),
    total_count = n(),
    pct_matched = matched_count / total_count * 100,
    .groups = "drop"
  )

@@

ggplot(summary_pct, aes(x = refClass, y = matched_count2, fill = refClass)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Percentage of Participants Matching Target Pct by Condition - 2% leeway",
    x = "Reference Class",
    y = "Percentage Matched (%)",
    fill = "Rounding"
  ) +
  theme_minimal()
